Reproducible Research
============================

*Course project 1*


**Loading and preprocessing the data**
1. Loading and preprocessing the data
  
  ```{r simulatedata}
  data<-read.csv("C:/Users/Anant Sharma/Desktop/DATA SCIENCE/Reproducible Research/Course project 1/activity.csv")
  ```
  
2. Processing the data:
  ```{r processdata}
  dim(data)
  names(data)
  str(data)
  ```

**What is mean total number of steps taken per day?**

1. Total number of steps taken per day:
  ```{r totalsteps}
  q1<-data.frame(total.steps=tapply(data$steps,data$date,sum,na.rm=TRUE))
  q1$date<-rownames(q1)
  rownames(q1)<-NULL
  q1
  
  ```
  
2. Histogram of the total number of steps taken each day:
  ```{r histogram}
  q2<-data.frame(sum=tapply(data$steps,data$date,sum,na.rm=TRUE))
  q2$date<-rownames(q2)
  rownames(q2)<-NULL
  library(ggplot2)
  g<-ggplot(q2,aes(x=date,y=sum))
  png(file="plot1.png")
  g+geom_bar(stat="identity")+xlab("date")+ylab("Total steps")+ggtitle("Total steps on every day")
  dev.off()
  g+geom_bar(stat="identity")+xlab("date")+ylab("Total steps")+ggtitle("Total steps on every day")
  png(file="plot1.1.png")
  hist(q2$sum,xlab="Total steps",ylab="Count",main="Total steps histogram",breaks=25)
  dev.off()
  hist(q2$sum,xlab="Total steps",ylab="Count",main="Total steps histogram",breaks=25)
  
  ```
3. Mean and Median of the total number of steps taken per day:
  ```{r mean}
  q3<-data.frame(mean.steps=tapply(data$steps,data$date,mean))
  q3$date<-rownames(q3)
  rownames(q3)<-NULL
  q3$median<-with(data,tapply(data$steps,data$date,median))
  q3
  ```
**What is the average daily activity pattern?**

1. Time series plot of the average number of steps taken:
  ```{r avgsteps}
  q4<-data.frame(mean.steps=tapply(data$steps,data$date,mean,na.rm=TRUE))
  q4$date<-rownames(q4)
  library(lubridate)
  q4$date<-ymd(q4$date)
  rownames(q4)<-NULL
  library(ggplot2)
  g<-ggplot(q4,aes(x=date,y=mean.steps))
  png(file="plot2.png")
  g+geom_bar(stat="identity")+xlab("Date")+ylab("average steps")+ggtitle("Time series plot")
  dev.off()
  g+geom_bar(stat="identity")+xlab("Date")+ylab("average steps")+ggtitle("Time series plot")
  ```
  
2. 5 minute interval that contains maximum steps on average:
  ```{r maximum}
  q5<-data.frame(mean=tapply(data$steps,data$interval,mean,na.rm=TRUE))
  q5$interval<-rownames(q5)
  rownames(q5)<-NULL
  q5[max(q5$mean),2]
  ```
  
**Imputing missing values**

1. Total number of missing values:
  ```{r missing}
  sum(is.na(data$steps))
  ```

2. Strategy for filling in all of the missing values in the dataset and creating a new dataset:
  ```{r strat}
  newdata<-data
  newdata$steps[is.na(newdata$steps)]<-mean(newdata$steps,na.rm=TRUE)
  ```
3. Histogram of the total number of steps taken each day for new data:
  ```{r newhist}
  newsub<-data.frame(sum=tapply(newdata$steps,newdata$date,sum))
  newsub$date<-rownames(newsub)
  rownames(newsub)<-NULL
  png(file="plot10.png")
  hist(newsub$sum,xlab="Total steps",ylab="count",main="Total number of steps",breaks=25)
  dev.off()
  hist(newsub$sum,xlab="Total steps",ylab="count",main="Total number of steps",breaks=25)
  ```
  
  
  *The NA values which were not plotted before are now plotted and the graph is changed because of it.*
  
4. Mean and Median of total number of steps taken per day
  ```{r mean2}
  q6<-data.frame(mean.steps=tapply(newdata$steps,newdata$date,mean))
  q6$date<-rownames(q6)
  rownames(q6)<-NULL
  q6$median<-with(newdata,tapply(newdata$steps,newdata$date,median))
  q6
  ```
  
  
  *The mean and median differ in the both the date.*
  
  **Are there differences in activity patterns between weekdays and weekends?**
  
  1. Factor variable in the dataset with two levels – “weekday” and “weekend”:
  ```{r week,echo=TRUE}
    library(lubridate)
    day<-weekdays(ymd(newdata$date))
    logic<-day %in%  c("Friday","Monday","Thursday","Tuesday","Wednesday")
    day[logic]<-"weekdays"
    day[!logic]<-"weekends"
    newdata$days<-day
  ```
  
  2. Average number of steps taken graph:
  ```{r graph}
      sub1<-newdata[newdata$days=="weekdays",]
      sub2<-newdata[newdata$days=="weekends",]
      sub11<-data.frame(mean=tapply(sub1$steps,sub1$interval,mean))
      sub11$interval<-rownames(sub11)
      rownames(sub11)<-NULL
      sub22<-data.frame(mean=tapply(sub2$steps,sub2$interval,mean))
      sub22$interval<-rownames(sub22)
      rownames(sub22)<-NULL
      sub11<-sub11[c("interval","mean")]
      sub22<-sub22[c("interval","mean")]
      png(file="plot20.png")
      plot(sub11,type="l",col="red",xlab="interval",ylab="avg steps",main="Average steps wrt to interval")
      lines(sub22,col="blue")
      legend("topleft",col=c("red","blue"),legend=c("weekdays","weekends"),lty=1:1,title="line type")
      dev.off()
      plot(sub11,type="l",col="red",xlab="interval",ylab="avg steps",main="Average steps wrt to interval")
      lines(sub22,col="blue")
      legend("topleft",col=c("red","blue"),legend=c("weekdays","weekends"),lty=1:1,title="line type")
      
  ```
  
  
  3. Panel plot:
  ```{r panel}
      png(file="plot30.png")
      par(mfrow=c(2,1))
      plot(sub11,type="l",xlab="interval",ylab="avg steps",main="weekdays")
      plot(sub22,type="l",xlab="interval",ylab="avg steps",main="weekends")
      dev.off()
      par(mfrow=c(2,1))
      plot(sub11,type="l",xlab="interval",ylab="avg steps",main="weekdays")
      plot(sub22,type="l",xlab="interval",ylab="avg steps",main="weekends")
      
  ```    
  